UZH IT \& AI Emissions Analysis\vspace{-1em}

### 2025-01-29 it-emissions.tex
Various model improvements: - more accurate user counts - clearer labels - PUE for on-campus data centers and cloud - streamlined code for usage and embodied emissions - training overhead / embodied emissions for LLMs

By contrast, \textcite{wikiuzh} reports approximately 28,000 students and 10,000 staff
at UZH. Evidently the spreadsheet only lists university assets, not personal devices.

\assumption{In modeling overall university IT emissions as well as potential LLM use,
we assume that the entire university population has a laptop and a mobile phone.}
We do not yet have any information about the breakdown of different functions / job roles, so
\assumption{we assume that IT use is uniform across the university}. Both of these assumptions
should be revisited.

@@

The best estimates are that use of large language model (LLM) like ChatGPT averages 3-4 Wh / request,
or approximately 10x the energy of a traditional search engine query. This number appears to be decreasing
rapidly as model hardware and software improve, even though overall energy use is increasing due to
increased utilization.

Nevertheless, for the model we assume that queries consume 3 Wh each, and we model per-capita daily
queries using a truncated normal distribution with mean 15 and $\sigma = 5$.

@@

>>>>

By contrast, \textcite{uzhprofile} reports approximately 28,000 students and 7,300 staff
at UZH. Evidently the spreadsheet only lists university assets, not personal devices.

\assumption{In modeling overall university IT emissions as well as potential LLM use,
we assume that 90\% of the university population, 32,000 people, have a laptop and a mobile phone.}
We do not yet have any information about the breakdown of different functions / job roles, so
\assumption{we assume that IT use is uniform across the university}. Both of these assumptions
should be revisited.

@@

The best estimates are that use of large language model (LLM) like ChatGPT averages 3-4 Wh / request,
or approximately 10x the energy of a traditional search engine query. This number appears to be decreasing
rapidly as model hardware and software improve, even though overall energy use is increasing due to
increased utilization. Nevertheless, for the model we assume that queries consume 3 Wh each.
We assume that on average every user will issue 15 queries per day.

@@

The embodied emissions and training emissions of LLMs are difficult to estimate and attribute precisely.
However, \textcite{devries2023growing} provides the following information about ChatGPT:
\begin{itemize}
    \item Operational power consumption of 500 MWh / day.
    \item Approximately 1300 MWh consumed in training.
    \item Approximately 4000 servers.
\end{itemize}

Assuming a relatively clean grid at 200 g CO$_2$e / kWh (a low estimate), and that each server has embodied emissions of 4 t CO$_2$e (a high estimate),
we obtain:
\begin{itemize}
    \item 100 t CO$_2$e / day operational emissions.
    \item 260 t CO$_2$e training emissions.
    \item 16 kt CO$_2$e total embodied emissions.
\end{itemize}

After one year of operation, operational emissions are approximately 36 kt CO$_2$e, whereas training and embodied emissions are about 16 kt CO$_2$e,
so operational emissions account for $\approx 70\%$ of the total. With less conservative assumptions (dirtier grid, lower embodied emissions),
the proportion of operational emissions increases further.

As a result, when modeling overall LLM emissions, we multiply estimated operational emissions by 1.4 to account for training and embodied emissions.
We acknowledge that this is, at best, a rough estimate based on just one LLM.
